---
title: "Simulations"
subtitle: "USC R Boot Camp: Statistical Computing"
author: "Emil Hvitfeldt"
date: "2019-08-19"
output:
  xaringan::moon_reader:
    css: ["default", "theme.css", "colors.css", "text-sizing.css"]
    lib_dir: libs
    nature:
      beforeInit: "macros.js"
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: [center, middle]
---

```{r include=FALSE}
library(knitr)
hook_output <- knit_hooks$get("output")
knit_hooks$set(output = function(x, options) {
  lines <- options$output.lines
  if (is.null(lines)) {
    return(hook_output(x, options))  # pass to default hook
  }
  x <- unlist(strsplit(x, "\n"))
  more <- "..."
  if (length(lines) == 1) {        # first n lines
    if (length(x) > lines) {
      # truncate the output, but add ....
      x <- c(head(x, lines), more)
    }
  } else {
    x <- c(more, x[lines], more)
  }
  # paste these lines together
  x <- paste(c(x, ""), collapse = "\n")
  hook_output(x, options)
})
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})

opts_chunk$set(
  echo = TRUE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618, # 1 / phi
  out.width = "700px")
```

# Statistical distributions

R has a very nice set of families of distributions. In general, distribution functions have the following name structure:

- Random Number Generation: r[name-of-the-distribution], e.g. `rnorm()` for normal, `runif()` for uniform.
- Density function: d[name-of-the-distribution], e.g. `dnorm()` for normal, `dunif()` for uniform.
- Cumulative Distribution Function (CDF): p[name-of-the-distribution], e.g. `pnorm()` for normal, `punif()` for uniform.
- Inverse of the CDF (quantile): q[name-of-the-distribution], e.g. `qnorm()` for the normal, `qunif()` for the uniform

---

# Statistical distributions

```{r}
rnorm(10)
rnorm(10, mean = 10)
```

See `?Distributions` for more details

---

# Setting the seed

```{r}
set.seed(1234)
rnorm(10)
```

```{r}
set.seed(1234)
rnorm(10)
```

---

# Sample

```{r}
dice <- c(1, 2, 3, 4, 5, 6)

sample(dice)

sample(dice, 1)

sample(dice, 100, replace = TRUE)
```

---

# Graphical interface

```{r, eval=FALSE}
normals <- rnorm(100)

plot(normals)

hist(normals)
```

---

# Tabulate

```{r}
dice <- 1:6

throws <- sample(dice, 100, replace = TRUE)

table(throws)
```

---

# Birthday problem

> If you have n people, what is the properbility that at least two people have the same birthday?

- What is the probability for n = 20?
- (hard) How many people do you need for the properbilty to be 50%?

Hint:

use `unique()`

---

# Simulate pi

- Sample a random point within a square going from -1 to 1 for both x and y.
- Count if point lies within inscribed circle. (Distence between point and (0,0 ) is less then 1.)
- Repeat

The procentage of hits should approximate pi/4.

hint: use `runif()` and Pythagoras' Theorem 

---

# Monty hall problem

> Suppose you're on a game show, and you're given the choice of three doors: Behind one door is a car; behind the others, goats. You pick a door, say No. 1, and the host, who knows what's behind the doors, opens another door, say No. 3, which has a goat. He then says to you, "Do you want to pick door No. 2?" Is it to your advantage to switch your choice?

- How often do you win if you always pick the first door and don't switch.
- (Hard) Write a version where you are allowed to switch.
